Creatinine Reverification
========================================================
This program implements calculations and plotting commands for annual reverification of creatinine screening on the Hitachi 717 automated chemistry analyzer.  
 
**Author** David Watson  
**email** [dewatson@icloud.com](mailto:dewatson@icloud.com)  
**Copyright** (c) 2014, David Watson  
  
Analysis began:
```{r timestamp began, echo = FALSE, comment = ""}
timestamp()
```
Reverification of LoQ, ULoL
---
```{r common utilities, echo = FALSE}
# Load custom SD, CV, and carryover functions; configure plots.
source("bin/common.R")
library("ggplot2")
```
```{r load cR data, echo = FALSE}
# Read in creatinine data and register column names:
cRdata <- read.csv("data/cR.csv",
                   header = TRUE,
                   colClasses = 
                     c("character","factor","numeric"))
attach(cRdata)
```
### Sample table
```{r check cR data, echo = FALSE, comment = ""}
cRdata
```
#### Mean, SD, and CV for each level
```{r statistics, echo = FALSE, comment = ""}
crMeans <- aggregate(Response, by = list(ID), mean)
colnames(crMeans) <- c("ID", "mean")
crStdDev <- aggregate(Response, by = list(ID), stddev)
colnames(crStdDev) <- c("ID", "SD")
crCV <- aggregate(Response, by = list(ID), cv)
colnames(crCV) <- c("ID", "CV")
merge(merge(crMeans, crStdDev), crCV)
```
```{r cleanup cR data, echo = FALSE}
#Unregister the cR data:
detach(cRdata)
```
---
Blank carryover analysis
---
```{r set decision point low carryover, echo = FALSE}
DecisionPoint <- 2.0
```
**Note: the decision point for this assay is `r DecisionPoint`**
```{r load blank carryover data, echo = FALSE}
# Read in carryover data:
lowCOdata <- read.csv("data/cRlowCO.csv",
                      header = TRUE,
                      colClasses =
                        c("character","character","numeric","logical"))
```
### Sample table
```{r verify blank carryover data, echo = FALSE, comment = ""}
lowCOdata
```
```{r blank carryover treatments, echo = FALSE}
# Select Armbruster "normal" and "carryover candidate" samples:
D_normal <- subset(lowCOdata, Candidate == FALSE, select = Response)
colnames(D_normal) <- c("D_N")
attach(D_normal)
D_candidate <- subset(lowCOdata, Candidate == TRUE, select = Response)
colnames(D_candidate) <- c("D_cc")
attach(D_candidate)
```
Carryover (%):
```{r print low carryover, echo = FALSE, comment = ""}
write(carryover(D_cc, D_N), "")
```
CV (%) of normal decision point samples: 
```{r print low CV, echo = FALSE, comment = ""}
write(round(100 * sd(D_N)/ mean(D_N), 1), "")
```
An unpaired, two-sample *t*-test using unequal variance (Welch) is used to check
whether the difference in means is < `r 0.2 * DecisionPoint` (20% of the decision point)
```{r blank carryover t-test, echo = FALSE, comment = ""}
t.test(D_N,
       D_cc,
       mu = 0.2 * DecisionPoint,
       alternative="less")
detach(D_normal)
detach(D_candidate)
```
---
High carryover analysis
---
```{r set decision point high carryover, echo = FALSE}
DecisionPoint <- 2.0
```
**Note: the decision point for this assay is `r DecisionPoint`**
```{r load high carryover data, echo = FALSE}
highCOdata <- read.csv('data/cRhighCO.csv',
                       header = TRUE,
                       colClasses = 
                         c("character","character","numeric","logical"))
```
### Sample table
```{r verify high carryover data, echo = FALSE, comment = ""}
highCOdata
```
```{r assign high carryover treatments, echo = FALSE}
#Select Armbruster "normal" and "carryover candidate" samples:
D_normal <- subset(highCOdata, Candidate == FALSE, select = Response)
colnames(D_normal) <- c("D_N")
attach(D_normal)
D_candidate <- subset(highCOdata, Candidate == TRUE, select = Response)
colnames(D_candidate) <- c("D_cc")
attach(D_candidate)
```
Carryover (%): 
```{r print high carryover, echo = FALSE, comment = ""}
write(carryover(D_cc, D_N), "")
```
CV (%) of normal decision point samples:
```{r print high CV, echo = FALSE, comment = ""}
write(round(100 * sd(D_N)/ mean(D_N), 1), "")
```
An unpaired, two-sample *t*-test using unequal variance (Welch) is used to check whether the difference in means is < `r 0.2 * DecisionPoint` (20% of the decision point)
```{r high carryover t-test, echo = FALSE, comment = ""}
t.test(D_cc,
       D_N,
       mu = 0.2 * DecisionPoint,
       alternative = "less")
detach(D_normal)
detach(D_candidate)
```
---
Cross-reactivity analysis
---
### Sample Table
```{r load cross-reactivity data, echo = FALSE, comment = ""}
crossrx <- read.csv('data/cRCR.csv',
                    header = TRUE)
crossrx
```
### Ascorbic acid cross-reactivity
```{r calculate ascorbic acid cross-reactivity, echo = FALSE}
ascorbicacid <- subset(crossrx,
                       Analyte == "ascorbate",
                       select = c(Concentration, Response))
ascorbicacid$Concentration <- ascorbicacid$Concentration * 100
attach(ascorbicacid)
aa_fit <- lm(Response ~ Concentration)
aa_summary <- summary(aa_fit)
aa_cross_reactivity <-
  100 * 20 * coef(aa_fit)[2]/(20.0 - coef(aa_fit)[1])
```
Ascorbic acid cross-reactivity (%):
```{r print ascorbic acid cross-reactivity, echo = FALSE, comment = ""}
write(round(as.numeric(aa_cross_reactivity), 1), "")
```
Ascorbic acid linear regression coefficients:
```{r ascorbic acid regression and plot, echo = FALSE, comment = "", warning = FALSE, fig.height = 6, fig.width = 6, dpi = 300}
aa_fit$coefficients
AAplot <- ggplot(ascorbicacid, aes(x=Concentration, y=Response, xmin = 0, ymin = 0))
AAplot + 
  geom_point(size=4) + 
  theme_classic(base_size = 18, base_family="Myriad Pro") + 
  geom_abline(intercept = coef(aa_fit)[1], slope = coef(aa_fit)[2], size=2, alpha=0.4) + 
  geom_abline(intercept = 2, slope = 0, linetype=3) +
  geom_abline(intercept = 20, slope = 0, linetype=3) +  
  labs(title="Ascorbic acid cross-reactivity", x="Concentration (mg/dL)", y="Response (mg/dL)") + 
  theme(plot.title = element_text(family = "Myriad Pro Semibold")) +
  annotate("text", x = Concentration[3], y = 0, 
           label=paste(
             "RSD:  ", 
             round(aa_summary$sigma, digits = 5), "\n",
             "r:  ", round(sqrt(aa_summary$r.squared), digits = 5), "\n",
             "m:  ", round(coef(aa_fit)[2], digits = 5), "\n",
             "b:  ", round(coef(aa_fit)[1], digits = 5),
             sep=""
           ),
           hjust=1, vjust=0)
detach(ascorbicacid)
```
### Niacin cross-reactivity
```{r calculate niacin cross-reactivity, echo = FALSE}
niacin <- subset(crossrx,
                 Analyte == "niacin",
                 select = c(Concentration, Response))
niacin$Concentration <- niacin$Concentration * 100
attach(niacin)
niacin_fit <- lm(Response ~ Concentration)
niacin_summary <- summary(niacin_fit)
niacin_cross_reactivity <-
  100 * 20 * coef(niacin_fit)[2]/(20.0 - coef(niacin_fit)[1])
```
Niacin cross-reactivity (%):
```{r print niacin cross-reactivity, echo = FALSE, comment = ""}
write(round(as.numeric(niacin_cross_reactivity), 1), "")
```  
Niacin linear regression coefficients:
```{r niacin regression and plot, echo = FALSE, comment = "", warning = FALSE, fig.width = 6, fig.height = 6, dpi = 300}
niacin_fit$coefficients
Niacinplot <- ggplot(niacin, aes(x=Concentration, y=Response, xmin = 0, ymin = 0))
Niacinplot + 
  geom_point(size=4) + 
  theme_classic(base_size = 18, base_family="Myriad Pro") + 
  geom_abline(intercept = coef(niacin_fit)[1], slope = coef(niacin_fit)[2], size=2, alpha=0.4) + 
  geom_abline(intercept = 2, slope = 0, linetype=3) +
  geom_abline(intercept = 20, slope = 0, linetype=3) +  
  labs(title="Niacin cross-reactivity", x="Concentration (mg/dL)", y="Response (mg/dL)") + 
  theme(plot.title = element_text(family = "Myriad Pro Semibold")) +
  annotate("text", x = Concentration[3], y = 0, 
           label=paste(
             "RSD:  ", 
             round(niacin_summary$sigma, digits = 5), "\n",
             "r:  ", round(sqrt(niacin_summary$r.squared), digits = 5), "\n",
             "m:  ", round(coef(niacin_fit)[2], digits = 5), "\n",
             "b:  ", round(coef(niacin_fit)[1], digits = 5),
             sep=""
           ),
           hjust=1, vjust=0)
detach(niacin)
```
Analysis finished:
```{r timestamp finished, echo = FALSE, comment = ""}  
timestamp()
```
```{r save cross-reactivity data, echo = FALSE}
# Save cross-reactivity data for plots:
save(ascorbicacid,
     aa_fit,
     aa_summary,
     niacin,
     niacin_fit,
     niacin_summary,
     file="data/cRCR.Rdata"
)
```
